---
title: "Correlation between Vazym and IDT samples at different spike-in ratios, min.cov = `r params$min.cov`"
author:
  - "hieunguyen@genesolutions.vn"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: no
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: true
    theme: lumen
params:
  min.cov: NA
---


```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

<style type="text/css">
    div.datatables { height: auto !important;}
</style>


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=14, fig.height=10}
# gc()
# rm(list = ls())

# min.cov <- 5
min.cov <- params$min.cov
analysis.version <- "0.1"
sample.type <- "all"

library(ggpubr)
path.to.main.src <- "/home/hieunguyen/PBMC_methylKit/src_v2"
source(file.path(path.to.main.src, "import_libraries.R"))
source(file.path(path.to.main.src, "helper_functions.R"))
source(file.path(path.to.main.src, "configs.R"))

min.cov.bases <- configs[[analysis.version]]$min.cov.bases
qvalue.cutoff <- configs[[analysis.version]]$qvalue.cutoff
methdiff.cutoff <- configs[[analysis.version]]$methdiff.cutoff
log2FC.cutoff <- configs[[analysis.version]]$log2FC.cutoff
up.flank <- configs[[analysis.version]]$up.flank
down.flank <- configs[[analysis.version]]$down.flank

maindir <- "/home/hieunguyen/PBMC_methylKit"
path.to.input <- file.path(maindir, "input_5x")
path.to.main.output <- file.path(maindir, "output_v2")

path.to.04.output <- file.path(path.to.main.output, sprintf("04_output_%s", analysis.version), sprintf("minCov_%s_sampleType_%s", min.cov, sample.type))
dir.create(path.to.04.output, showWarnings = FALSE, recursive = TRUE)

original.all.cov.files <- Sys.glob(file.path(path.to.input, sprintf("filtered_%sreads_cov", min.cov), "*.cov"))
names(original.all.cov.files) <- unlist(lapply(original.all.cov.files, function(x){
  x <- basename(x)
  x <- str_replace(x, ".deduplicated.bedGraph.gz.bismark.zero.filtered.cov", "")
  # x <- str_split(x, "-")[[1]][[2]]
  # x <- str_split(x, "_")[[1]][[1]]
  return(x)
}))

meta.data <- data.frame(filename = names(original.all.cov.files))
meta.data <- meta.data %>% rowwise() %>%
  mutate(Sample.type = ifelse(grepl("BCI", filename) == TRUE, "I", "V")) %>%
  mutate(spike.in.Label = str_split(filename, "_")[[1]][[1]]) %>%
  mutate(spike.in.Label = str_split(spike.in.Label, "-")[[1]][[2]]) %>%
  mutate(spike.in.Label = str_replace(spike.in.Label, "BCI", "")) %>%
  mutate(spike.in.Label = str_replace(spike.in.Label, "BCV", "")) %>%
  mutate(spike.in.Label = str_split(spike.in.Label, "R")[[1]][[1]])

original.metadata <- meta.data 

meth <- hash()
mat <- hash()
mat.pivot <- hash()

for (chosen.spike.in.label in unique(original.metadata$spike.in.Label)){
  meta.data <- subset(original.metadata, original.metadata$spike.in.Label %in% c(chosen.spike.in.label))
  meta.data <- meta.data %>% rowwise() %>%
    mutate(Label = ifelse(Sample.type == "I", 1, 0))
  all.cov.files <- original.all.cov.files[meta.data$filename]
  labels <- meta.data$Label
  
  DML.obj <- readBismarkCoverage( all.cov.files,
                       sample.id = names(all.cov.files),
                       assembly = "hg19",
                       treatment = labels,
                       context = "CpG",
                       min.cov = min.cov)
  meth[[chosen.spike.in.label]] <- methylKit::unite(object = DML.obj, destrand = FALSE) 
  mat[[chosen.spike.in.label]] <- percMethylation(meth[[chosen.spike.in.label]], rowids = TRUE)  %>% as.data.frame() %>% rownames_to_column("locus")
  mat.pivot[[chosen.spike.in.label]] <- mat[[chosen.spike.in.label]] %>% pivot_longer(!locus, names_to = "sample", values_to = "meth_density")
  mat.pivot[[chosen.spike.in.label]] <- merge(mat.pivot[[chosen.spike.in.label]], meta.data, by.x = "sample", by.y = "filename", all.x = TRUE)
}

```

# Correlation plot between sample I and sample V at the same spike-in ratio {.tabset}

## Spike-in ratio 0.1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.keep='all', fig.width=14, fig.height=10, }
chosen.spike.in.label <- "0.1"
a <- getCorrelation(meth[[chosen.spike.in.label]], plot = TRUE)
```

## Spike-in ratio 1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.keep='all', fig.width=14, fig.height=10, }
chosen.spike.in.label <- "01"
a <- getCorrelation(meth[[chosen.spike.in.label]], plot = TRUE)
```

## Spike-in ratio 5
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.keep='all', fig.width=14, fig.height=10, }
chosen.spike.in.label <- "05"
a <- getCorrelation(meth[[chosen.spike.in.label]], plot = TRUE)
```

## Spike-in ratio 10
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.keep='all', fig.width=14, fig.height=10, }
chosen.spike.in.label <- "10"
a <- getCorrelation(meth[[chosen.spike.in.label]], plot = TRUE)
```

## Spike-in ratio 100
```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide',fig.keep='all', fig.width=14, fig.height=10, }
chosen.spike.in.label <- "100"
a <- getCorrelation(meth[[chosen.spike.in.label]], plot = TRUE)
```

# Boxplot: Comparison of I samples and V samples at same spike-in ratios {.tabset}

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10, }
for (group in names(mat)){
  cat(sprintf("## Spike-in ratio: %s \n", group))
  p <- mat.pivot[[group]] %>% ggplot(aes(x = Sample.type, y = meth_density, fill = Sample.type)) + 
  geom_boxplot() + stat_compare_means(method = "t.test", comparisons = list(c("I", "V"))) +
  theme_pubr(base_size = 25) + ggtitle(sprintf("I samples and V samples at spike-in ratio %s", group))
  print(p)
  cat("\n \n")
}
```

# Violin plot: Comparison of I samples and V samples at same spike-in ratios {.tabset}
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10, }
for (group in names(mat)){
  cat(sprintf("## Spike-in ratio: %s \n", group))
  p <- mat.pivot[[group]] %>% ggplot(aes(x = Sample.type, y = meth_density, fill = Sample.type)) + 
  geom_violin() + stat_compare_means(method = "t.test", comparisons = list(c("I", "V"))) +
  theme_pubr(base_size = 25) + ggtitle(sprintf("I samples and V samples at spike-in ratio %s", group))
  print(p)
  cat("\n \n")
}

```

# Correlation between different spike-in ratios and MEAN methylation density in I samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat <- data.frame()
for (group in names(mat.pivot)){
  tmp.mat <- rbind(tmp.mat, subset(mat.pivot[[group]], mat.pivot[[group]]$Sample.type == "I"))
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat %>% ggplot(aes(x = spike.in.Label, y = meth_density)) + geom_boxplot()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat %>% ggplot(aes(x = spike.in.Label, y = meth_density)) + geom_violin()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
convert.ratio <- list(`0` = 0, 
                      `0.1` = 0.1,
                      `01` = 1,
                      `05` = 5,
                      `10` = 10,
                      `100` = 100)
mean.table <- tmp.mat %>% group_by(spike.in.Label) %>%  summarise_at(vars(meth_density), list(name = mean)) %>%
  rowwise() %>%
  mutate(spike.in.ratio = convert.ratio[[spike.in.Label]])

mean.table %>%
  ggplot(aes(x = log2(spike.in.ratio), y = name)) + geom_point() +  geom_smooth() + 
  scale_x_continuous(breaks = log2(sort(mean.table$spike.in.ratio)), labels = mean.table$spike.in.ratio) +
  stat_cor(method = "pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 14) + theme_pubr(base_size = 25)
  
  

```


# Correlation between different spike-in ratios and MEAN methylation density in V samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat <- data.frame()
for (group in names(mat.pivot)){
  tmp.mat <- rbind(tmp.mat, subset(mat.pivot[[group]], mat.pivot[[group]]$Sample.type == "V"))
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat %>% ggplot(aes(x = spike.in.Label, y = meth_density)) + geom_boxplot()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
mean.table <- tmp.mat %>% group_by(spike.in.Label) %>%  summarise_at(vars(meth_density), list(name = mean)) %>%
  rowwise() %>%
  mutate(spike.in.ratio = convert.ratio[[spike.in.Label]])

mean.table %>%
  ggplot(aes(x = log2(spike.in.ratio), y = name)) + geom_point() +  geom_smooth() + 
  scale_x_continuous(breaks = log2(sort(mean.table$spike.in.ratio)), labels = mean.table$spike.in.ratio) +
  stat_cor(method = "pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 14)+ theme_pubr(base_size = 25)
  
  

```

# Correlation between different spike-in ratios and median methylation density in I samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat <- data.frame()
for (group in names(mat.pivot)){
  tmp.mat <- rbind(tmp.mat, subset(mat.pivot[[group]], mat.pivot[[group]]$Sample.type == "I"))
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat %>% ggplot(aes(x = spike.in.Label, y = meth_density)) + geom_boxplot()
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
convert.ratio <- list(`0` = 0, 
                      `0.1` = 0.1,
                      `01` = 1,
                      `05` = 5,
                      `10` = 10,
                      `100` = 100)
median.table <- tmp.mat %>% group_by(spike.in.Label) %>%  summarise_at(vars(meth_density), list(name = median)) %>%
  rowwise() %>%
  mutate(spike.in.ratio = convert.ratio[[spike.in.Label]])

median.table %>%
  ggplot(aes(x = log2(spike.in.ratio), y = name)) + geom_point() +  geom_smooth() + 
  scale_x_continuous(breaks = log2(sort(median.table$spike.in.ratio)), labels = median.table$spike.in.ratio) +
  stat_cor(method = "pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 14) + theme_pubr(base_size = 25)



```


# Correlation between different spike-in ratios and median methylation density in V samples
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat <- data.frame()
for (group in names(mat.pivot)){
  tmp.mat <- rbind(tmp.mat, subset(mat.pivot[[group]], mat.pivot[[group]]$Sample.type == "V"))
}
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
tmp.mat %>% ggplot(aes(x = spike.in.Label, y = meth_density)) + geom_boxplot()
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
median.table <- tmp.mat %>% group_by(spike.in.Label) %>%  summarise_at(vars(meth_density), list(name = median)) %>%
  rowwise() %>%
  mutate(spike.in.ratio = convert.ratio[[spike.in.Label]])

median.table %>%
  ggplot(aes(x = log2(spike.in.ratio), y = name)) + geom_point() +  geom_smooth() + 
  scale_x_continuous(breaks = log2(sort(median.table$spike.in.ratio)), labels = median.table$spike.in.ratio) +
  stat_cor(method = "pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 14)+ theme_pubr(base_size = 25)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
all.DML.obj <- readBismarkCoverage( original.all.cov.files,
                       sample.id = names(original.all.cov.files),
                       assembly = "hg19",
                       treatment = to_vec( for (item in original.metadata$spike.in.Label) if (item == "0") 0 else 1),
                       context = "CpG",
                       min.cov = min.cov)

all.meth <- methylKit::unite(object = all.DML.obj, destrand = FALSE)

all.mat <- hash()
for (i in seq(1, length(all.DML.obj))){
  tmpdf <- data.frame(all.DML.obj[[i]]@.Data)
  colnames(tmpdf) <- all.DML.obj[[i]]@names
  all.mat[[all.DML.obj[[i]]@sample.id]] <- tmpdf
}

count.CpG.in.sample <- data.frame(filename = names(all.mat))
count.CpG.in.sample <- merge(count.CpG.in.sample, original.metadata, by.x = "filename", by.y = "filename") %>%
  rowwise() %>%
  mutate(countCpG = all.mat[[filename]] %>% nrow())
```

# Number of CpG in each sample
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
count.CpG.in.sample %>% ggplot(aes(x = filename, y = countCpG, fill = Sample.type)) + geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle= 90, vjust = 0.5, hjust = 0.5))
```

# Correlation all samples V and I
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
sample.order <- original.metadata %>% arrange(desc(Sample.type))
sample.order <- sample.order$filename
meth.mat <- percMethylation(all.meth)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
print(sprintf("Number of CpG existing in all 36 samples: %s", nrow(meth.mat)))
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
cor.mat <- cor(meth.mat) %>% as.data.frame()
cor.mat <- cor.mat[sample.order, sample.order]

pheatmap::pheatmap(cor.mat, cluster_rows = FALSE, cluster_cols = FALSE, scale = "none", show_rownames = TRUE)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
##### HELPER FUNCTIONS
plot_corr_and_mse_heatmap <- function(spike.in.ratio1, spike.in.ratio2, sample.type){
  library(RColorBrewer)
  breaksList <- seq(0.7, 1, by = 0.01)
  input.samples <- subset(original.metadata, original.metadata$Sample.type == sample.type 
                          & original.metadata$spike.in.Label %in% 
                            c(spike.in.ratio1, spike.in.ratio2))$filename
  
  input.mat <- meth.mat[, input.samples]
  tmp.cor.mat <- cor(input.mat) %>% as.data.frame()
  
  corr.heatmap <- pheatmap::pheatmap(tmp.cor.mat, 
                                     cluster_rows = FALSE, 
                                     cluster_cols = FALSE, 
                                     scale = "none", 
                                     show_rownames = TRUE, 
                                     display_numbers = TRUE, 
                                     fontsize_number = 20, 
                                     fontsize_row = 20, 
                                     fontsize_col = 20, 
                                     color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)), 
                                     breaks = breaksList)
  
  mse.mat <- data.frame(col = colnames(input.mat))
  for (x in unique(mse.mat$col)){
    mse.mat[[x]] <- unlist(lapply(mse.mat$col, function(y){
      input.sample1 <- x
      input.sample2 <- y
      
      tmp.mat <- input.mat[, c(input.sample1, input.sample2)] %>% as.data.frame()
      colnames(tmp.mat) <- c("V1", "V2")
      tmp.mat$sq.diff <- (tmp.mat$V1 - tmp.mat$V2)^2
      return(sqrt(sum(tmp.mat$sq.diff)))
    }))
  }

  
  mse.heatmap <- pheatmap::pheatmap(mse.mat %>% column_to_rownames("col"), 
                                    cluster_rows = FALSE, 
                                    cluster_cols = FALSE, 
                                    scale = "none", 
                                    show_rownames = TRUE, 
                                    display_numbers = TRUE, 
                                    fontsize_number = 20, 
                                    fontsize_row = 20, fontsize_col = 20)
  return(list(cor = corr.heatmap, mse = mse.heatmap))
}
```

# Correlation at each pair of spike-in ratio in sample V
## Spike-in ratio 0 vs 0.1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "0"
ratio2 <- "0.1"
input.sampletype <- "V"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 0.1 vs 1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "0.1"
ratio2 <- "01"
input.sampletype <- "V"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 1 vs 5
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "01"
ratio2 <- "05"
input.sampletype <- "V"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 5 vs 10
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "05"
ratio2 <- "10"
input.sampletype <- "V"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```


## Spike-in ratio 10 vs 100
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "10"
ratio2 <- "100"
input.sampletype <- "V"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

# Correlation at each pair of spike-in ratio in sample I

## Spike-in ratio 0 vs 0.1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "0"
ratio2 <- "0.1"
input.sampletype <- "I"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 0.1 vs 1
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "0.1"
ratio2 <- "01"
input.sampletype <- "I"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 1 vs 5
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "01"
ratio2 <- "05"
input.sampletype <- "I"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

## Spike-in ratio 5 vs 10
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "05"
ratio2 <- "10"
input.sampletype <- "I"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```


## Spike-in ratio 10 vs 100
```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
ratio1 <- "10"
ratio2 <- "100"
input.sampletype <- "I"
output <- plot_corr_and_mse_heatmap(spike.in.ratio1 = ratio1,
                                    spike.in.ratio2 = ratio2,
                                    sample.type = input.sampletype)
```

# Correlation at each CpG site


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
corrdf <- data.frame()
for (i in seq(1, nrow(meth.mat))){
  tmp.methdf <- data.frame(meth = meth.mat[i, ]) %>% rownames_to_column("filename") 
  tmp.methdf <- merge(tmp.methdf, original.metadata, by.x = "filename", by.y = "filename")
  
  tmp.methdf <- tmp.methdf %>% rowwise() %>%
    mutate(spike.in.Label = convert.ratio[[spike.in.Label]])
  
  tmp.methdf.I <- subset(tmp.methdf, tmp.methdf$Sample.type == "I")
  tmp.methdf.V <- subset(tmp.methdf, tmp.methdf$Sample.type == "V")
  
  tmp.cor.I <- cor(tmp.methdf.I$meth, tmp.methdf.I$spike.in.Label)
  tmp.cor.V <- cor(tmp.methdf.V$meth, tmp.methdf.V$spike.in.Label)
  
  tmp.outputdf <- data.frame(CpG = c(i), pearson.cor.I = c(tmp.cor.I), pearson.cor.V = c(tmp.cor.V))
  corrdf <- rbind(corrdf, tmp.outputdf)  
}
corrdf.pivot <- corrdf %>% pivot_longer(!CpG, names_to = "sample.type", values_to = "corr")
corrdf.pivot %>% ggplot(aes(x = sample.type, y = corr)) + 
  geom_boxplot() + stat_compare_means(size = 10, label.x = 1.1, label.y = 0) + theme_pubr(base_size = 25)
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
print(sprintf("Mean Pearson correlation in V samples: %s", mean(na.omit(corrdf$pearson.cor.V))))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=14, fig.height=10}
print(sprintf("Mean Pearson correlation in I samples: %s", mean(na.omit(corrdf$pearson.cor.I))))
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='asis',fig.keep='all', fig.width=15, fig.height=15}
# input.sample1 <- input.samples[[1]]
# input.sample2 <- input.samples[[4]]
# 
# tmp.mat <- input.mat[, c(input.sample1, input.sample2)] %>% as.data.frame()
# colnames(tmp.mat) <- c("V1", "V2")
# tmp.mat$sq.diff <- (tmp.mat$V1 - tmp.mat$V2)^2
# 
# tmp.plot <- tmp.mat  %>%
#   ggplot(aes(x = V1, y = V2)) + 
#   geom_point(alpha = 0.1, size = 4) +
#   stat_cor(method = "pearson", p.accuracy = 0.001, r.accuracy = 0.01, size = 14) +
#   theme(title = element_text(size = 20),
#         axis.title = element_text(size = 20),
#         axis.text.x = element_text(size = 20),
#         axis.text.y = element_text(size = 20)) +
#   ggtitle(sprintf("Mean square error: %s", round(sqrt(sum(tmp.mat$sq.diff))))) +
#   xlab(input.sample1) + ylab(input.sample2) +
#   xlim(c(0, 100)) + ylim(c(0, 100))
# 
# tmp.plot
```